---
title: Alcohol consumption and happiness
author: Marcell Puski
date: '2021-05-29'
slug: alcohol-consumption-and-happiness
categories: []
tags:
  - R Markdown
  - plot
---


In this short analysis i was curious about the connection of countries state of global happiness and their alcohol consumptions. In countries where people are generally satisfied and happy overall less alcohol is consumed? Or is it the opposit, that in countries where people tend to drink more, are people also happier?  

To answer this question i am investigating two datasets available online in [kagle](https://www.kaggle.com/).

One of the datasets is the [World Happiness Report](https://www.kaggle.com/unsdsn/world-happiness?select=2015.csv) which is a landmark survey of the state of global happiness. It contains 158 countries estimated happiness score (10 - happiest 0 -least happy) and the ranking based on that score. It also containes six factors: economic production, social support, life expectancy, freedom, absence of corruption, and generosity, which represents the contribution of this factors of making life evaluations higher in each country than they are in Dystopia, a hypothetical country that has values equal to the world’s lowest national averages for each of the six factors.
I used the *World Happiness Report 2015*.   


The other dataset is the [Alcohol Comsumption around the World](https://www.kaggle.com/codebreaker619/alcohol-comsumption-around-the-world) which contains information of 193 countries alcohol consumptions, provided by the *World Health Organisation, Global Information System on Alcohol and Health (GISAH), 2010*. The dataset is providing information about on avarage how many glasses of wine, cans of beer and shots of spirits and the total liters of pure alcohol were drunk per person in each country in 2010. 

I combined this two datasets together so i got a dataset that contains information's about the countries happiness leveles and alcohol consumption as well. Unfortunatelly there is five years difference in the data collection for the happiness scores and alcohol consumption, but probably these characteristics of countries citisens changes slowly, and overall it only causes minor inaccuracies.  


```{r include=FALSE}
#  loading libraries

library(tidyverse)
library(readxl)
library(skimr)
library(dplyr)
library(knitr)

```


```{r include=FALSE}
# loading data

world_happiness <- read_csv("data/2015.csv")

alcohol <- read.csv("data/drinks.csv")

```


```{r include=FALSE}

# joining data

# renaming country column to be the same in the two datasets
world_happiness <- 
  world_happiness %>%
  rename(country = 'Country')

# joining the two datasets based on country names
dt <- left_join(world_happiness, alcohol)


# checking joined data
apply(is.na(dt),2,sum)

# creating data set for missing data
missing_dt <- 
  dt %>%
  filter(is.na(beer_servings))

```



```{r include=FALSE}

# in a few cases there is a mismatch between the datasets in the countries names causing more missing data
# The following names represent the same country:
    # United States = USA
    #Russia = Russian Federation
    #Congo (Brazzaville) = Congo
    #Congo (Kinshasa) = DR Congo
    #Trinidad and Tobago = Trinidad & Tobago
    #Somaliland region	= Somalia
    #Bosnia and Herzegovina = Bosnia-Herzegovina
    #Ivory Coast = Cote d'Ivoire

# correcting country names 

# creating coppy for safty
chumpa <- world_happiness 


# list of country names which needs to be replaced

csere <- c("United States", "Trinidad and Tobago", "Russia", "Somaliland region", "Bosnia and Herzegovina", "Congo (Brazzaville)","Congo (Kinshasa)","Ivory Coast")


# list of corrected country names

bere <- c("USA", "Trinidad & Tobago", "Russian Federation", "Somalia", "Bosnia-Herzegovina", "Congo", "DR Congo", "Cote d'Ivoire")


# renaming country names
szerint <- which(chumpa$country %in% csere)

for (i in 1:length(szerint)) {
  chumpa$country[szerint[i]] <-  bere[i]
  
  
}

# re joining data, with corrected names
dt_clear <- left_join(chumpa, alcohol)


# checking 
dim(dt_clear)
apply(is.na(dt_clear),2,sum)


```



```{r include=FALSE}
# Deleting rows where there is no data about alcohol consumtion 

dt_clear <- na.omit(dt_clear)

# checking
dim(dt_clear)

apply(is.na(dt_clear),2,sum)

```



```{r include=FALSE}
# renaming columns without spaces

dt_clear <- 
  dt_clear %>%
  rename(happiness_score = 'Happiness Score',
         happiness_rank = 'Happiness Rank',
         standard_error = 'Standard Error',
         GDP = 'Economy (GDP per Capita)',
         life_expectancy = 'Health (Life Expectancy)',
         corruption = 'Trust (Government Corruption)',
         dystopia_residual = 'Dystopia Residual',
         total_alcohol = 'total_litres_of_pure_alcohol')

```



```{r include=FALSE}
# creating sub-groups for alcohol consumption

dt_clear$alcohol_cat <- cut(dt_clear$total_alcohol, c(-1, 0, 5,10,15), labels = c("doesn't drink","drinks less than 5 liters", "drinks 5-10 liters", "drinks 10-15 liters"))

```

By joining the datasets together and keeping only the countries where is informations about alcohol drinking and happiness as well, there remained 153 countries. 


Table 1 shows the five happiest countries, and Table 2 shows the top five most alcohol drinkinng countries.  

```{r echo=FALSE, message=FALSE}
# table of 10 happiest countries

dt_clear %>%
  select(country:happiness_score, beer_servings:total_alcohol) %>%
  filter(happiness_rank <= 5) %>%
  kable(., digits = 2, align = "l", col.names = c("Country", "Region", "Happiness Rank", "Happiness Score", "Beer servings", "Spirit servings", "Wine servings", "Total liters of pure alcohol"), caption = "Ten five countries")


```



```{r echo=FALSE, message=FALSE}

# table of 10 most alcohol drinking countries
dt_clear %>%
  select(country:happiness_score, beer_servings:total_alcohol) %>%
  arrange(desc(total_alcohol)) %>%
  top_n(5) %>%
  kable(., digits = 2, align = "l", col.names = c("Country", "Region", "Happiness Rank", "Happiness Score", "Beer servings", "Spirit servings", "Wine servings", "Total liters of pure alcohol"), caption = "Top five countries that drinks the most alcohol")


```



# Descriptive statistics by region's

```{r echo=FALSE}



# avarages about happiness and alcohol consumtion


# creating summary table by Regions - usufull but doesn't look good

# dt_clear %>%
#   group_by(Region) %>%
#   summarise(
#     mn_happiness = mean(happiness_score),
#     sd_happiness = sd(happiness_score),
#     max_happiness = max(happiness_score),
#     min_happiness = min(happiness_score),
#     mn_alcohol = mean(total_alcohol),
#     sd_alcohol = sd(total_alcohol),
#     max_alcohol = max(total_alcohol),
#     min_alcohol = min(total_alcohol)
#   ) %>%
#   kable(., digits = 2, align = "l", col.names = c("Region", "mean happiness", "happiness standard deviation", "max happiness score", "minimum happiness score", "mean total liters of alcohol", "standard deviation of total liters of alcohol", "maximum total liters of alcohol", "minimum total liters of alcohol"), caption = "Summary table by Regions")

# Ploting summary table

plot_1 <- dt_clear %>%
  group_by(Region) %>%
  summarise(
    mn_happiness = mean(happiness_score),
    sd_happiness = sd(happiness_score),
    max_happiness = max(happiness_score),
    min_happiness = min(happiness_score),
    mn_alcohol = mean(total_alcohol),
    sd_alcohol = sd(total_alcohol),
    max_alcohol = max(total_alcohol),
    min_alcohol = min(total_alcohol)
  )



plot_1 %>%
  ggplot() +
  geom_bar(aes(x = Region, y = mn_happiness), stat = "identity", fill = "skyblue1") +
  geom_errorbar(aes(x = Region, ymin = mn_happiness - sd_happiness, ymax = mn_happiness + sd_happiness), color = "red3") +
  labs(title = "Happiness levels by regions",
       x = "", y = "mean happiness score") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90))

plot_1 %>%
  ggplot() +
  geom_bar(aes(x = Region, y = mn_alcohol), stat = "identity", fill = "skyblue1") +
  geom_errorbar(aes(x = Region, ymin = mn_alcohol - sd_alcohol, ymax = mn_alcohol + sd_alcohol), color = "red3") +
  labs(title = "Alcohol consumption by regions",
       x = "", y = "mean total liters of alcohol") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90))
  

# alcohol / happiness rank + Region

dt_clear %>%
  arrange(happiness_rank) %>%
  ggplot() +
  geom_point(mapping = aes(x = happiness_rank, y = total_alcohol, color = Region)) +
  labs(title = "Happiness rank and alcohol consumption by region's",
       x = "Happiness rank",
       y = "mean total liters of alcohol") +
  theme_classic()



# # alcohol /happinass rank + Region - Doesn't look good 
# dt_clear %>%
#   arrange(happiness_rank) %>%
#   ggplot() +
#   geom_point(mapping = aes(x = happiness_rank, y = total_alcohol, color = Region)) +
#   facet_wrap(~Region) +
#   labs(title = "Happiness rank and alcohol consumption by region's",
#        x = "Happiness rank",
#        y = "mean total liters of alcohol") +
#   theme_classic()
# 


```






# Does happy countries drink less alcohol?

In the following let's investigate whether more alcohol consumption in a country also goes together with less happy citizens. 
Obviusly it is not possible to draw too powerfull conclusions from this investigation, while there are too many athor factors that could influence this alcohol - happiness connection, but it would be interesting to see, how this two factors goes together with.  



```{r echo=FALSE, message=FALSE, warning=FALSE}

# alcohol + happiness rank plot 

dt_clear %>%
  arrange(happiness_rank) %>%
  ggplot(., mapping = aes(x = happiness_rank, y = total_alcohol)) +
  geom_point(shape = 21, color = "black", fill = "steelblue", alpha = 1/2, size = 2) +
  geom_smooth(color = "firebrick") +
  labs(title = "Happiness rank and alcohol consumption",
       x = "Happiness rank",
       y = "mean total liters of alcohol") +
  theme_classic()


dt_clear %>%
  arrange(happiness_rank) %>%
  ggplot(., mapping = aes(x = happiness_score, y = total_alcohol)) +
  geom_point(shape = 21, color = "black", fill = "steelblue", alpha = 1/2, size = 2) +
  geom_smooth(color = "firebrick") +
  geom_hline(yintercept = mean(dt_clear$total_alcohol), linetype="dotted", color = "black", size = 1,5) +
  labs(title = "Happiness score and alcohol consumption",
       x = "Happiness score",
       y = "mean total liters of alcohol") +
  theme_classic()




```


It seems like that exacly the opposit relationship is observable from the data, that the more alcohol is consumed in a country, the happier it's citizens are.  


In order to further investigate i created categories based on the amount of total alcohol consumption. 


```{r echo=FALSE}

dt_clear %>%
  group_by(alcohol_cat) %>%
  summarise(
    mean_happiness = mean(happiness_score),
    sd_happiness = sd(happiness_score),
    mean_happy_rank = mean(happiness_rank),
    sd_happy_rank = sd(happiness_rank)
  ) %>%
  kable(., digits = 2, align = "l", col.names = c("Alcohol category", "mean happiness", "happiness standard deviation", "mean happiness rank", "happiness rank standard deviation"), caption = "Avarage happiness levels by alcohol consumption categories")


dt_clear %>%
  group_by(alcohol_cat) %>%
  summarise(
    mean_happiness = mean(happiness_score),
    sd_happiness = sd(happiness_score),
    mean_happy_rank = mean(happiness_rank),
    sd_happy_rank = sd(happiness_rank)
  ) %>%
  ggplot() +
  geom_bar(aes(x = alcohol_cat, y = mean_happiness), stat = "identity", fill = "skyblue1") +
  geom_hline(yintercept = mean(dt_clear$happiness_score), linetype="dotted", color = "red", size = 2) +
  labs(title = "Alcohol categories and mean happiness scores",
       x = "",
       y = "mean happiness scores") +
  theme_classic()



```

From this combined data set it seems like in happy countries people tend to drink more. Obviusly there could not be drawn causation from this data, it is not clear weather alcohol makes people more happy, or if they are more happy they tend to drink more alcohol. 
Also there could be other factors like cultural differences in alcohol consumption that could affect this relationship. 



# Countries where people do not drink alcohol

In the following let's see only the countries where people don't drink alcohol


```{r echo=FALSE}
# summary table 
dt_clear %>%
  filter(total_alcohol == 0) %>%
  select(happiness_rank, country, happiness_score, total_alcohol) %>%
  kable(., digits = 2, align = "l", col.names = c("Happiness rank", "Country", "Happiness score", "total liters of pure alcohol"), caption = "Countries where people doesn't drink alcohol")



# plot
dt_clear %>%
  filter(total_alcohol == 0) %>%
  arrange(happiness_rank) %>%
  ggplot() +
  geom_bar(mapping = aes(x = Region, y = happiness_score, fill = country), stat = "identity", position = "dodge") +
  geom_hline(yintercept = mean(dt_clear$happiness_score), linetype="dotted", color = "red") +
  labs(title = "Countries where people doesn't drink alcohol",
       x = "",
       y = "mean happiness scores") +
  theme_classic()


```



# Do people in happy countries drink different kind of alcohol than people in less happy countries?


In the present dataset there is information about how much servings of an avarage person drinks beer, spirits, and wine in a country, so it would be interesting to see weather people in happy countries has different alcohol drinking habbits than is less happy countries.  


For this reason i also created subcategories now based on the countries happines ranks, where the first 50 countries i labeled *"happy"* countries between 50 and 100 *"medium happy"* countries and above 100 *"not happy"* countries.

This is of course a subjective category only for demonstration purposes.

```{r echo=FALSE}
# for safety
plot_servings <- dt_clear

# creating happiness categories

plot_servings$happy_cat <- cut(plot_servings$happiness_rank, c(-1, 50, 100, 160), labels = c("happy", "medium happy", "not happy"))



# table

plot_servings %>%
  group_by(happy_cat) %>%
  summarise(
    mn_beer = mean(beer_servings),
    mn_spirit = mean(spirit_servings),
    mn_wine = mean(wine_servings),
    mn_total = mean(total_alcohol)
  ) %>%
  kable(., digits = 2, align = "l", col.names = c("happyness category", "mean beer servings", "mean spirit servings", "mean wine servings", "mean total liters of alcohol"), caption = "types of alcohol servings by happiness")


# make the data long format to be able to plot it

plot_servings_plot <- plot_servings %>%
  group_by(happy_cat) %>%
  summarise(
    mn_beer = mean(beer_servings),
    mn_spirit = mean(spirit_servings),
    mn_wine = mean(wine_servings),
    mn_total = mean(total_alcohol)
  ) %>%
  gather(data = ., 
       key = alcohol_type, value = servings,
       'mn_beer':'mn_wine')


# plot

plot_servings_plot %>%
  ggplot() +
  geom_bar(mapping = aes(x = happy_cat, y = servings, fill = alcohol_type), stat = "identity", position = "dodge") +
  labs(title = "Types of alcohol servings by happiness",
       x = "",
       y = "mean servings",
       fill = "alcohol type") +
  theme_classic()

```


One difference between *happy* and *medium* and *not happy* countries is the wine consumption. In every category people drink beer the most, but in the top 50 rated country people drink as much wine as spirits, while in less happy countries people drink more spirits than wine.



# question three - lets see sub categories of happiness: GDP etc.



```{r echo=FALSE, message=FALSE}
# GDP

dt_clear %>%
  ggplot(., mapping = aes(x = GDP, y = total_alcohol)) +
  geom_point(shape = 21, color = "black", fill = "steelblue", alpha = 1/2, size = 2) +
  geom_smooth(color = "firebrick") +
  ylim(0, 15)

# Family
dt_clear %>%
  ggplot(., mapping = aes(x = Family, y = total_alcohol)) +
  geom_point(shape = 21, color = "black", fill = "steelblue", alpha = 1/2, size = 2) +
  geom_smooth(color = "firebrick") +
  ylim(0, 15)

# life_expectansy

dt_clear %>%
  ggplot(., mapping = aes(x = life_expectancy, y = total_alcohol)) +
  geom_point(shape = 21, color = "black", fill = "steelblue", alpha = 1/2, size = 2) +
  geom_smooth(color = "firebrick") +
  ylim(0, 15)

# Freedom

dt_clear %>%
  ggplot(., mapping = aes(x = Freedom, y = total_alcohol)) +
  geom_point(shape = 21, color = "black", fill = "steelblue", alpha = 1/2, size = 2) +
  geom_smooth(color = "firebrick") +
  ylim(0, 15)

# corruption

dt_clear %>%
  ggplot(., mapping = aes(x = corruption, y = total_alcohol)) +
  geom_point(shape = 21, color = "black", fill = "steelblue", alpha = 1/2, size = 2) +
  geom_smooth(color = "firebrick") +
  ylim(0, 15)

# generosity

dt_clear %>%
  ggplot(., mapping = aes(x = Generosity, y = total_alcohol)) +
  geom_point(shape = 21, color = "black", fill = "steelblue", alpha = 1/2, size = 2) +
  geom_smooth(color = "firebrick") +
  ylim(0, 15)


```


